### 2021-04-19 Chapter-3/Chapter-3.tex
Updates from Overleaf

Figure~\ref{fig:memory_hierarchy} shows the overview of one example of the hierarchy memory architecture. For a CPU with multiple cores, each core has two private level-1 (L1) cache, an instruction cache (iCache), and a data cache (dCache). Because instructions and data have different access patterns, separating iCache and dCache makes it possible for CPUs to fetch the code and data and improve performance simultaneously. Followed by the level-1 cache, each core also has the unified private level-2 cache for both the code and the data. The level-3 cache is the last level cache (LLC). The LLC is shared among all the cores. The main memory is in the bottom-most under the memory hierarchy.
\begin{figure}
  \centering
  \includegraphics[width=.5\columnwidth]{./figures/chapter3/architecture.pdf}
  \caption{Computer memory hierarchy. The hierarchy is designed to minimise the access time with the relatively low cost. It is the root cause of many side-channel attacks.}\label{fig:memory_hierarchy}
\end{figure}

The current computer memory hierarchy opens the way for side-channel attacks from two aspects. First, the architecture relies on the system software for managing the memory, which becomes a problem in a threat model where the operating system is not untrusted (e.g., Intel SGX). Second, the size of the cache is smaller than the main memory. It is possible that different units in the main memory share the same address in the cache.
\begin{figure}
  \centering
  \includegraphics[width=.65\columnwidth]{./figures/chapter3/address.pdf}
  \caption{Memory addressing}\label{fig:memory_address}
\end{figure}

@@

\subsection{Imprecision}
In this dissertation, we study two types of side-channel leakage code patterns. The intuition is that the target application shows different control flows or data flows when it processes different sensitive input data. We refer them as \textit{secret-dependent control flow transfers} and \textit{secret-dependent data accesses}. Unlike previous work, our tool works on the machine code, which is more precise than previous tools.
\subsubsection{Control Flow}
\begin{figure}[h]
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
int example1(uint8_t k, uint32_t m) { // m = 0
  uint32_t b = (k & m) >> 7;
  if (b) {
    ...    // Branch 1
  } else {
    ...    // Branch 2
  }
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
push  ebp
mov   ebp, esp
movzx eax, [addr_k]    
and   eax, [addr_me] 
shr   eax, 0x7           
test  eax, eax
jne   branch 2
...
\end{lstlisting}
  \end{minipage}\caption*{(a) A False Negative}

  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
int example2(uint16_t k) {
  int res;
  if (k > 8) {
    res = 0;
  } else {
    res = 1;
  }
  return res;
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
xor   eax, eax
cmp   [addr_k], 8
setbe  al
xor    edx, edx
ret
\end{lstlisting}
  \end{minipage}\caption*{(b) A False Positive}
  \caption{Secret-dependent Control-flow Transfers}\label{fig:chapter3:cf}
\end{figure}

If the input-sensitive data can affect the victim program's control flow, then an attacker can infer the sensitive data by observing the control flow during the execution. Figure~\ref{fig:chapter3:cf}(a) shows such an example. The function takes a public value \textsf{m} and a secret \textsf{k} as the input values. The code at line 2 ensures that the lower 7 bits of the value \textsf{k} do not affect the value \textsf{b}. However, depending on the value of \text{m}, the code snippet in Figure~\ref{fig:chapter3:cf}(a) may have a leakage site. If the eighth bit of \textsf{m} is 1, then an attacker can also infer the eighth bit of \textsf{k} by observing the branch at line 4 or line 6. The right figure shows the corresponding machine code. It has a conditional jump instruction \textsf{jne}. The sensitive input \textsf{k} can affect the program counter (the rip register), which is the root cause of the side-channel leakages.

Figure~\ref{fig:chapter3:cf}(b) shows a different situation. While the sensitive input \textsf{k} can affect the if-else branch, the compiler removes the branch with a conditional set instruction.  The opposite situation is the same. For example, recent work tries to use bit masking to rewrite the program with control flows. However, for some situations~\cite{Coppens:2009:PMT:1607723.1608124}, compilers (e.g., GCC) optimize the code too much (from a security view) and remove the unnecessary copy by reintroducing conditional moves.
\subsubsection{Data Flow}
\begin{figure}[h]
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[128];
...
int example3(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 128;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x7f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
  \end{minipage}\caption*{(a) A True Leakage}

  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[32];
...
int example4(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 32;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x3f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
  \end{minipage}\caption*{(b) A False Positive}
  \caption{Secret-dependent Data Accesses}\label{fig:chapter3:da}
\end{figure}

@@

\subsection{Performance}
The second limitation of the current side-channel detection tool is the performance bottleneck, especially for those tools based on the symbolic analysis. For example, CaSym can only analyze small programs. CacheD perfroms better. However, it still can not handle large programs like RSA directly. It uses some domain knowledge to analyze only part of the program. The expensive cost comes from the symbolic execution. Symbolic execution can be used to explore all the possible paths, which is useful for many tasks. But it is significantly less likely for cache side channels bugs in crypto primitives. Cryptography primitives are more likely to have an even coverage of different paths (based on pseudo-random cipher states), and a branch-based side channel is vulnerable from the very first branch on secret data, making complex path conditions less often relevant. Therefore, we choose to analyze a trace instead of the whole path. Another bottleneck is that the IR. While the IR can significantly decrease the difficulty of the implementation, it also gives introduce an average of 10x of the overall overhead. Third, using SMT solving can be convenient because of its generality, and it is commonly used for its precision when checking branch feasibility in symbolic execution. But the decision problems that an SMT solver handles are typically at least NP-hard. On the other hand, we can use some methods to avoid the solver and find side-channel leakages in a more efficient way.

\section{Method}\label{chapter3:method}
\begin{figure}
  \centering
  \includegraphics[width=.3\columnwidth]{./figures/chapter3/attack.pdf}
  \caption{A side-channel attack}\label{fig:side-channel-attack}
\end{figure}

@@

For one execution, a program ($\beta$) has many temporary values ($t_i \in
  T$). Once $\beta$ (program), $k$ (secret), and $m$ (message, public) are
determined, $t_i$ is also fixed. Therefore, $ t_i = f_i(\beta, k, m)$, where $f_
  i$ is a function that maps between $t_i$ and ($\beta$, $k$, $m$). In the paper,
we consider two code patterns that can be exploited by an attacker,
\emph{secret-dependent control transfers} and \emph{secret-dependent data
  accesses}. In other words, an adversary has observations based on control-flows
and data accesses.

\subsection{Secret-dependent Control Transfers}
A control-flow path is secret-dependent if different input-sensitive keys
($K$) can lead to different branch conditions.
We define a branch to be secret-dependent if:
$$\exists k_{i1}, k_{i2} \in K, m \in M, \,f_i(\beta, k_{i1}, m) \neq f_i(\beta, k_{i2}, m)$$

An adversary can observe which branch the code executes if the branch condition
equals $t_b$. We use the constraint $c_i : f_i(\beta, k, m) = t_b$ to model
the observation ($o$) on secret-dependent control-transfers. Note that in the
above definition, $m$ is also a variable. So it is possible for some $m \in M$,
we can not find a pair of two different $k_{i1}$ and $k_{i2}$ to satisfy the above
inequation. For example, in the example from Figure~\ref{fig:chapter3:cf}, if $m = 0$,
the code will always execute the branch 2. However, the function \textsf{example1} still
has a secret-dependent control-flow for some $m$, so we think the code has
some side-channel leakages.

@@

$$\exists k_{i1}, k_{i2} \in K, m \in M,\,f_i(\beta, k_{i1}, m) >> L \neq f_i(\beta, k_{i2}, m) >> L$$

@@

\begin{table}%[ht]
  \centering\small\footnotesize
  \caption{The number of x86,  % instructions and the number of 
    REIL IR, and VEX IR instructions on the traces of crypto programs.}
  \label{scala:ir}
  \resizebox{.7\columnwidth}{!}{%

@@

The tool runs symbolic execution on top of execution traces. At the beginning of the symbolic execution, the tool creates new symbols for each byte in the raw buffer. For other data in the register or memory at the beginning, we use actual values from the runtime information collected during the runtime. During the symbolic execution for each instruction, the tool updates every variable in the memory and registers with a math formula. The formula is made up of concrete values and the input key as the symbols accumulated through the symbolic execution.
For each formula, the tool will check weather it can be reduced into a concrete values (e.g., $k_1+12-k_1 = 12$ ). If so, the tool will only use the concrete values in the following symbolic execution.

@@

\begin{lstlisting}[xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
 ...
0x0000e781      add dword [local_14h], 1
0x0000e785      cmp dword [local_14h], 4
0x0000e789      jne 0xe7df
0x0000e78b      mov dword [local_14h], 0
 ...
\end{lstlisting}

At the beginning of the instruction segment, the value at the address of local\_14h can be written as $F(\vec{K})$. At the address $e785$, the value will be updated with $F(\vec{K})+1$. Then the code compares
the value with 4 and use the result as a conditional jump. Based on the result, we can have the following formula:

@@

The formula, together with the memory address ($0xe789$) is store as a \textit{formula tuple (address, formula)}. Each formula tuple represents one leakage site.

\subsubsection{Secret-dependent Data Accesses}
Like input-dependent control-flow transfers, an adversary can also infer sensitive information from the data access pattern as well. We try to find this kind of leakages by checking every memory operand of the instruction. We generate the memory addressing
formulas. As discussed before, every symbols in the formula is the input key. If the formula does not contain any symbols, the memory access is independent
from the input sensitive information and will not leak any sensitive information according to our threat model. Otherwise, we will generate the constraint for
the memory addressing. We model the memory address with a symbolic formula $F(\vec{K})$. Because we also have the concrete value of the memory address $Addr1$.
Inspired by the work from~\cite{203878}, the formula can be written as:

$$F(\vec{K}) >> L = Addr1 >> L$$

@@

\subsubsection{Information Flow Check}
\detect{} is designed to help software developers find and understand the side-channel vulnerabilities. To ease the procedure of fixing the bug, we also track the information flow for each byte of the input
buffer.
The step can be seen as the multiple-tag taint analysis. With the help of the information from symbolic execution, we can implement a relatively simple but relatively precise information flow track.
At the beginning of the analysis, \detect{} keep a track for each byte in
the original buffer. When \detect{} symbolically executes each instruction in the trace, it will check every value read from registers or memory. If the value is concrete, it means the instruction has nothing to do with the original buffer.
If the value is a formula, it means the original information passes through the instruction. Since each byte in the sensitive buffer is represented as a symbol with a unique ID, \detect{} can know which byte in the origin buffer goes through the instruction.

@@

\begin{table*}[]

@@

The comparison result with CacheD~\cite{203878} is shown in Table~\ref{eval:cacheD}.
Note that one statement in the source code can be compiled into several machine instructions. So it is possible that one statement can have multiple leakage points. Under the circumstance, we think it is only a leakage.
We have confirmed that \detect{} can identify all the secret-dependent data access vulnerabilities reported by CacheD. In addition, \detect{} finds many new ones.
CacheD fails to detect some vulnerabilities for two
reasons. First, CacheD can only detect secret-dependent data access
vulnerabilities. \detect{} can detect secret-dependent control-flows as well.
Second, according to the CacheD paper, CacheD times out after 20 hours to process
asymmetric ciphers. CacheD applies some domain knowledge to simplify and speed up
the analysis.
While those optimizations do not introduce false positives, they may miss some
vulnerabilities.
Third, CacheD only symbolizes the encryption keys. Some side-channel leakages when the input messages equal to some certain values. It is possible that CacheD miss some
leakages as we have discussion the previous sections.
We notice that the number of instructions in those traces are different due to the different analysis starting functions and building options during the evaluation.
Table~\ref{eval:cacheD} shows that
\detect{} is faster than CacheD. \detect{} is much faster than CacheD when analyzing the same
number of instructions. For example, when we test~\detect{} on AES from OpenSSL
0.9.7, \detect{} is over 100x faster than CacheD.
\begin{table*}[]
 \centering
  \caption{Comparison with CacheD: Time,
    Secret-dependent Control-flows\,(CF), Secret-dependent Data-flows\,(DF), The Number of Instructions\,(\# Instructions).}
  \label{eval:cacheD}

    \begin{tabular}{@{}l|r@{~~}r@{~~}r|r@{~~}r@{~}r@{}}
      \hline
      \multicolumn{1}{l|}{} & \multicolumn{3}{c|}{CacheD} & \multicolumn{3}{c}{Abacus}                                             \\
      Name                  & Time (s)                    & \# Instructions            & DF & Time (s) & \# Instructions & CF+DF   \\ \hline
      AES OpenSSL 0.9.7         & 43.4                        & 791                        & 48 & 0.23     & 1,704           & 0+75    \\
      AES OpenSSL 1.0.2f       & 48.5                        & 2,410                      & 32 & 0.04     & 1,350           & 0+88    \\
      RSA OpenSSL 0.9.7       & 199.3                       & 674,797                    & 1  & 1,351.6  & 17,002,523      & 105+2   \\
      RSA OpenSSL 1.0.2f     & 165.6                       & 473,392                    & 1  & 1,753.3  & 14,468,307      & 27+11   \\
      RSA Libgcrypt 1.6.1     & 11542.3                     & 26,848,103                 & 2  & 128.1    & 9,527,231       & 9+2     \\
      RSA Libgcrypt 1.7.3         & 10788.9                     & 27,775,053                 & 0  & 891.7    & 10,513,606      & 14+0    \\ \hline
      Total                 & 22,788.0                    & 55,738,546                 & 84 & 4,125.0  & 51,514,721      & 155+178 \\ \hline
      \multicolumn{7}{l}{\# of Instructions per second \qquad  CacheD: 2,445 \qquad \tool: 12,489}                                 \\ \hline
    \end{tabular}

@@

DATA~\cite{217537} identifies
side-channel leakages by finding differences in execution traces of
the test program under \emph{various secret inputs}. According to
the original DATA paper, it uses 443 different traces to analyze the
side-channel vulnerabilities in symmetric cyphers and 450 different
traces to analyze the side-channel vulnerabilities in asymmetric
cyphers. On the other hand, \detect{} detects side-channel
vulnerabilities from one execution trace. \detect{} uses symbolic
analysis to extract formulas that model each side-channel
leakage. After that, we sample the formula with \emph{various
  secret inputs} to detect and quantify each leakage site. In
theory, DATA might have better code coverage than \detect{} because it
uses more execution traces,  but \detect{} has the following
advantages. a) \detect{} is faster than DATA. For example, it takes
116 minutes for DATA to detect vulnerabilities in the RSA implementation in OpenSSL 1.1.0f\@. \detect{} only spends 34 minutes, as shown in Table~\ref{table:over_result}. It takes 13 minutes and 20 minutes
for DATA to analyze the side-channel leakages in AES and DES,
respectively. On the other hand, \detect{} finishes its analysis in
less than ten seconds while finding all the leakages reported by
DATA. b) Because \detect{} does not exexute the test program again when we
have a new \emph{secret input}, \detect{} can test more input secrets
on those formulas within the same time to achieve better precision.
c) DATA tries to use leakage models (domain knowledge) to classify each leakage.
The strength of Abacus is that it does not need such domain knowledge.
DATA reports 278 control-flow and 460 memory-access leaks for the RSA implementation in OpenSSL 1.1.0f. Among those leakages, they find one new vulnerability in RSA after some manual analysis.
\detect{} finds the vulnerability.  For each leakage site, \detect{} can provide
concrete examples to trigger the issue and give an estimation to
assess the severity level of the vulnerability.

>>>>

Figure~\ref{fig:memory_hierarchy} shows the overview of one example of the hierarchy memory architecture. For a CPU with multiple cores, each core has two private level-1 (L1) cache, an instruction cache (iCache), and a data cache (dCache). Because instructions and data have different access patterns, separating iCache and dCache makes it possible for CPUs to fetch the code and data and improve performance simultaneously. Followed by the level-1 cache, each core also has the unified private level-2 cache for both the code and the data. The level-3 cache is the last level cache (LLC). The LLC is shared among all the cores. The main memory is in the bottom-most under the memory hierarchy.
\begin{figure}[ht]
  \centering
  \includegraphics[width=.5\columnwidth]{./figures/chapter3/architecture.pdf}
  \caption{Computer memory hierarchy. The hierarchy is designed to minimise the access time with the relatively low cost. It is the root cause of many side-channel attacks.}\label{fig:memory_hierarchy}
\end{figure}

The current computer memory hierarchy opens the way for side-channel attacks from two aspects. First, the architecture relies on the system software for managing the memory, which becomes a problem in a threat model where the operating system is not untrusted (e.g., Intel SGX). Second, the size of the cache is smaller than the main memory. It is possible that different units in the main memory share the same address in the cache.
\begin{figure}[ht]
  \centering
  \includegraphics[width=.65\columnwidth]{./figures/chapter3/address.pdf}
  \caption{Memory addressing}\label{fig:memory_address}
\end{figure}

@@

\subsection{Imprecision}
In this dissertation, we study two types of side-channel leakage code patterns. The intuition is that the target application shows different control flows or data flows when it processes different sensitive input data. We refer them as \textit{secret-dependent control flow transfers} and \textit{secret-dependent data accesses}. Unlike previous work, our tool works on the machine code, which is more precise than previous tools that work on the source code or the intermediate representation.
\subsubsection{Control Flow}
\begin{figure}[ht]
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
int example1(uint8_t k, uint32_t m) { // m = 0
  uint32_t b = (k & m) >> 7;
  if (b) {
    ...    // Branch 1
  } else {
    ...    // Branch 2
  }
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
push  ebp
mov   ebp, esp
movzx eax, [addr_k]    
and   eax, [addr_me] 
shr   eax, 0x7           
test  eax, eax
jne   branch 2
...
\end{lstlisting}
  \end{minipage}\caption*{(a) A False Negative}

  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
int example2(uint16_t k) {
  int res;
  if (k > 8) {
    res = 0;
  } else {
    res = 1;
  }
  return res;
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
xor   eax, eax
cmp   [addr_k], 8
setbe al
xor   edx, edx
ret
\end{lstlisting}
  \end{minipage}\caption*{(b) A False Positive}
  \caption{Secret-dependent control-flow transfers}\label{fig:chapter3:cf}
\end{figure}

If the input-sensitive data can affect the victim program's control flow, then an attacker can infer the sensitive data by observing the control flow during the execution. Figure~\ref{fig:chapter3:cf}(a) shows such an example. The function takes a public value \textsf{m} and a secret \textsf{k} as the inputs. The code at line 2 ensures that the lower 7 bits of the value \textsf{k} do not affect the value \textsf{b}. However, depending on the value of \text{m}, the code snippet in Figure~\ref{fig:chapter3:cf}(a) may have a leakage site. If the eighth bit of \textsf{m} is 1, then an attacker can infer the eighth bit of \textsf{k} by observing the branch at line 4 or line 6. The right figure shows the corresponding machine code. It has a conditional jump instruction \textsf{jne}. The sensitive input \textsf{k} can affect the program counter (the rip register), which is the root cause of the side-channel leakages.

However, the if-else branch does not always indicate there is a potential leakage site. Figure~\ref{fig:chapter3:cf}(b) shows a different situation. While the sensitive input \textsf{k} can affect the if-else branch, the compiler removes the branch with a conditional set instruction.  The opposite situation is the same. For example, recent work tries to use bit masking to rewrite the program with control flows. However, for some situations~\cite{Coppens:2009:PMT:1607723.1608124}, compilers (e.g., GCC) optimize the code too much (from a security view) and remove the unnecessary copy by reintroducing conditional moves.
\subsubsection{Data Flow}
\begin{figure}[ht]
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[128];
...
int example3(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 128;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x7f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
  \end{minipage}\caption*{(a) A True Leakage}

  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[32];
...
int example4(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 32;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\linewidth}
    \begin{lstlisting}[xleftmargin=.15\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x3f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
  \end{minipage}\caption*{(b) A False Positive}
  \caption{Secret-dependent data accesses}\label{fig:chapter3:da}
\end{figure}

@@

\subsection{Performance}
The second limitation of current side-channel detection tools is the performance bottleneck, especially for those tools based on the symbolic analysis. For example, CaSym can only analyze small programs. CacheD performs better than CaSym. However, it still can not handle large programs like RSA directly. It uses some domain knowledge to analyze only part of the program. The expensive cost comes from the symbolic execution. Symbolic execution can be used to explore all the possible paths, which is useful for many tasks.
Nevertheless, it is significantly less likely for cache side channel bugs in crypto primitives. Cryptography primitives are more likely to have an even coverage of different paths (based on pseudo-random cipher states), and a branch-based side channel is vulnerable from the very first branch on secret data, making complex path conditions less often relevant. Therefore, we choose to analyze a trace instead of the whole path. Another bottleneck is that the IR. While the IR can significantly decrease the difficulty of the implementation, it also gives introduce an average of 10x of the overall overhead. Third, using SMT solving can be convenient because of its generality, and it is commonly used for its precision when checking branch feasibility in symbolic execution. But the decision problems that an SMT solver handles are typically at least NP-hard. On the other hand, we can use some methods to avoid the solver and find side-channel leakages more efficiently.

\section{Method}\label{chapter3:method}
\begin{figure}[ht]
  \centering
  \includegraphics[width=.3\columnwidth]{./figures/chapter3/attack.pdf}
  \caption{A side-channel attack}\label{fig:side-channel-attack}
\end{figure}

@@

For one execution, a program ($\beta$) has many temporary values ($t_i \in
  T$). Once $\beta$ (program), $k$ (secret), and $m$ (message, public) are
determined, $t_i$ is also fixed. Therefore, $ t_i = f_i(\beta, k, m)$, where $f_
  i$ is a function that maps between $t_i$ and ($\beta$, $k$, $m$). In the paper,
we consider two code patterns that can be exploited by an attacker,
\emph{secret-dependent control transfers} and \emph{secret-dependent data
  accesses}.

\subsection{Secret-dependent Control Transfers}
A control-flow path is secret-dependent if different input-sensitive keys
($K$) can lead to different branch conditions.
We define a branch to be secret-dependent if:
$$\exists k_{i1}, k_{i2} \in K, m \in M, \,f_i(\beta, k_{i1}, m) \neq f_i(\beta, k_{i2}, m) \land PC$$

$PC$ represents the path condition. An adversary can observe which branch the code executes if the branch condition
equals $t_b$. We use the constraint $c_i : f_i(\beta, k, m) = t_b$ to model
the observation ($o$) on secret-dependent control-transfers. Note that in the
above definition, $m$ is also a variable. So it is possible for some $m \in M$,
we can not find a pair of two different $k_{i1}$ and $k_{i2}$ to satisfy the above inequation. For example, in Figure~\ref{fig:chapter3:cf}, if $m = 0$, the function \textsf{example1} will always execute the branch 2. However, if $m = \mathtt{0xffffffff}$ and $k = 0$, the function \textsf{example1} executes the branch 1 and the function \textsf{example1} executes the branch 2 when $k = \mathtt{0xf000000}$.

@@

$$\exists k_{i1}, k_{i2} \in K, m \in M,\,f_i(\beta, k_{i1}, m) >> L \neq f_i(\beta, k_{i2}, m) >> L \land PC$$

@@

\begin{table}[ht]
  \centering\small\footnotesize
  \caption{The number of x86,  % instructions and the number of 
    REIL IR, and VEX IR instructions on the traces of crypto programs.}
  \label{scala:ir}
  \resizebox{.7\columnwidth}{!}{%

@@

The tool runs symbolic execution on top of execution traces. At the beginning of the symbolic execution, the tool creates new symbols for each byte in the raw buffer. For other data in the register or memory at the beginning, we use actual values from the runtime information collected during the runtime. During the symbolic execution for each instruction, the tool updates every variable in the memory and registers with a math formula. The formula is made up of concrete values and the input key as the symbols accumulated through the symbolic execution. We implement several rules to simplify the formulas. For each formula, the tool will check weather it can be reduced into a concrete values (e.g., $k_1+12-k_1 = 12$ ). If so, the tool will only use the concrete values in the following symbolic execution.

@@

\begin{lstlisting}[xleftmargin=.2\textwidth, xrightmargin=.2\textwidth, frame=none][ht]
...
0x0000e781      add dword [local_14h], 1
0x0000e785      cmp dword [local_14h], 4
0x0000e789      jne 0xe7df
0x0000e78b      mov dword [local_14h], 0
...
\end{lstlisting}

At the address $\mathtt{0x0000e781}$, the value at the address of $\mathtt{local\_14h}$ can be written as $F(\vec{K})$. At the address $\mathtt{0x0000e785}$, the value will be updated with $F(\vec{K})+1$. Then the code compares the value with $4$ and use the result as a conditional jump. Based on the result, we can have the following formula:

@@

The formula, together with the memory address ($\mathtt{0x0000e789}$) is stored as a \textit{formula tuple (address, formula)}. Each formula tuple represents one potential leakage site.

\subsubsection{Secret-dependent Data Accesses}
Like input-dependent control-flow transfers, an adversary can also infer sensitive information from the data access pattern as well. We try to find this kind of leakages by checking every memory operand of the instruction. We generate the memory addressing
formulas. As discussed before, every symbols in the formula is the input key. If the formula does not contain any symbols, the memory access is independent
from the input sensitive information and will not leak any sensitive information according to our threat model. Otherwise, we will generate the constraint for
the memory addressing. We model the memory address with a symbolic formula $F(\vec{K})$. Because we also have the concrete value of the memory address $\mathtt{Addr1}$. Inspired by the work from~\cite{203878}, the formula can be written as:

$$F(\vec{K}) >> L = \mathtt{Addr1} >> L$$

@@

\subsubsection{Information Flow Check}
\detect{} is designed to help software developers find and understand the side-channel vulnerabilities. To ease the procedure of fixing the bug, we also track the information flow for each byte of the input
buffer.
The step can be seen as the multiple-tag taint analysis. With the help of the information from symbolic execution, we can implement a relatively simple but relatively precise information flow track.
At the beginning of the analysis, \detect{} keep track for each byte in
the original buffer. When \detect{} symbolically executes each instruction in the trace, it will check every value read from registers or memory. If the value is concrete, it means the instruction has nothing to do with the original buffer.
If the value is a formula, it means the original information passes through the instruction. Since each byte in the sensitive buffer is represented as a symbol with a unique ID, \detect{} can know which byte in the origin buffer goes through the instruction.

@@

\begin{table*}

@@

The comparison result with CacheD~\cite{203878} is shown in Table~\ref{eval:cacheD}.
Note that one statement in the source code can be compiled into several machine instructions. So it is possible that one statement can have multiple leakage points. Under the circumstance, we think it is only a leakage.
We have confirmed that \detect{} can identify all the secret-dependent data access vulnerabilities reported by CacheD. In addition, \detect{} finds many new ones.
CacheD fails to detect some vulnerabilities for two
reasons. First, CacheD can only detect secret-dependent data access
vulnerabilities. \detect{} can detect secret-dependent control-flows as well.
Second, according to the CacheD paper, CacheD times out after 20 hours to process
asymmetric ciphers. CacheD applies some domain knowledge to simplify and speed up
the analysis.
While those optimizations do not introduce false positives, they may miss some
vulnerabilities.
Third, CacheD only symbolizes the encryption keys. Some side-channel leakages when the input messages equal to some certain values. It is possible that CacheD miss some
leakages as we have discussion the previous sections.
We notice that the number of instructions in those traces are different due to the different analysis starting functions and building options during the evaluation.
Table~\ref{eval:cacheD} shows that
\detect{} is faster than CacheD. \detect{} is much faster than CacheD when analyzing the same
number of instructions. For example, when we test~\detect{} on AES from OpenSSL
0.9.7, \detect{} is over 100x faster than CacheD.
\begin{table*}[ht]
 \centering
  \caption{Comparison with CacheD: Time,
    Secret-dependent Control-flows\,(CF), Secret-dependent Data-flows\,(DF), The Number of Instructions\,(\# Instructions).}
  \label{eval:cacheD}

    \begin{tabular}{@{}l|r@{~~}r@{~~}r|r@{~~}r@{~}r@{}}
      \hline
      \multicolumn{1}{l|}{} & \multicolumn{3}{c|}{CacheD} & \multicolumn{3}{c}{Abacus}                                             \\
      Benchmark                  & Time (s)                    & \# Instructions            & DF & Time (s) & \# Instructions & CF+DF   \\ \hline
      AES OpenSSL 0.9.7         & 43.4                        & 791                        & 48 & 0.23     & 1,704           & 0+75    \\
      AES OpenSSL 1.0.2f       & 48.5                        & 2,410                      & 32 & 0.04     & 1,350           & 0+88    \\
      RSA OpenSSL 0.9.7       & 199.3                       & 674,797                    & 1  & 1,351.6  & 17,002,523      & 105+2   \\
      RSA OpenSSL 1.0.2f     & 165.6                       & 473,392                    & 1  & 1,753.3  & 14,468,307      & 27+11   \\
      RSA Libgcrypt 1.6.1     & 11542.3                     & 26,848,103                 & 2  & 128.1    & 9,527,231       & 9+2     \\
      RSA Libgcrypt 1.7.3         & 10788.9                     & 27,775,053                 & 0  & 891.7    & 10,513,606      & 14+0    \\ \hline
      Total                 & 22,788.0                    & 55,738,546                 & 84 & 4,125.0  & 51,514,721      & 155+178 \\ \hline
      \multicolumn{7}{l}{\# of Instructions per second \qquad  CacheD: 2,445 \qquad \tool: 12,489}                                 \\ \hline
    \end{tabular}

@@

DATA~\cite{217537} identifies
side-channel leakages by finding differences in execution traces of
the test program under \emph{various secret inputs}. According to
the original DATA paper, it uses 443 different traces to analyze the
side-channel vulnerabilities in symmetric cyphers and 450 different
traces to analyze the side-channel vulnerabilities in asymmetric
cyphers. On the other hand, \detect{} detects side-channel
vulnerabilities from one execution trace. \detect{} uses symbolic
analysis to extract formulas that model each side-channel
leakage. After that, we sample the formula with \emph{various
  secret inputs} to detect and quantify each leakage site. In
theory, DATA might have better code coverage than \detect{} because it
uses more execution traces,  but \detect{} has the following
advantages. a) \detect{} is faster than DATA. For example, it takes
116 minutes for DATA to detect vulnerabilities in the RSA implementation in OpenSSL 1.1.0f\@. \detect{} only spends 34 minutes, as shown in Table~\ref{table:over_result}. It takes 13 minutes and 20 minutes
for DATA to analyze the side-channel leakages in AES and DES,
respectively. On the other hand, \detect{} finishes its analysis in
less than ten seconds while finding all the leakages reported by
DATA. b) Because \detect{} does not exexute the test program again when we
have a new \emph{secret input}, \detect{} can test more input secrets
on those formulas within the same time to achieve better precision.
c) DATA tries to use leakage models (domain knowledge) to classify each leakage.
The strength of \ctool{} is that it does not need such domain knowledge.
DATA reports 278 control-flow and 460 memory-access leaks for the RSA implementation in OpenSSL 1.1.0f. Among those leakages, they find one new vulnerability in RSA after some manual analysis.
\detect{} finds the vulnerability.  For each leakage site, \detect{} can provide
concrete examples to trigger the issue.
\subsection{Case Studies}
\subsubsection{DES and AES}
Our tool confirms that both implementations in OpenSSL and mbedTLS have side-channel leakages
sites. Moreover, we find that all the leakages belong to the secret-dependent data accesses. The reference implementation of AES uses four lookup tables to speed up the computation. However, such implementation are vulnerable to side-channel attacks. Previous work has shown an end-to-end attack to fully recover the key. Our findings is similar to the previous work. We also notice that later version of OpenSSL (after 1.0.1) uses a modified version with smaller S-table. However, \detect{} also find the leakage sites in the version. Recent work also finds the sites through theoretical analysis.
